### 2022-03-08 ehoh/ehoh.tex
more

An easy but inefficient way to implement unification and matching for \lfhol{}
is to apply the applicative encoding (Sect.~\ref{sec:ehoh:logic}), perform
first-order unification or matching, and decode the resulting
substitution. To avoid the overhead, we generalize the first-order unification
and matching procedures to operate directly on \lfhol{} terms.

@@

The key difference between $\Gfpf$ and $\GfpfRTL$ concerns applied variable.
Given the sample positions $(\varepsilon, 2, 1)$,
the fingerprint of $x$ is $(\AA, \BB, \BB)$ as before, whereas
the fingerprint of $x \; \cst{c}$ is $(\AA, \BB, \cst{c}).$
%
As another example,
let $(\varepsilon, 2,\allowbreak 1,\allowbreak 2.2,\allowbreak 2.1,\allowbreak 1.2, 1.1)$
be the sample positions,
and let $s = x \; (\cst{f} \; \cst{b} \; \cst{c})$ and $t = \cst{g}\; \cst{a}\; (y \; \cst{d}).$
Their fingerprints are
$\FpRTL(s) = (\AA, \BB, \cst{f}, \BB, \BB, \cst{b}, \cst{c})$ and
$\FpRTL(t) = (\cst{g}, \cst{a}, \AA, \NN, \NN, \BB, \cst{d})$.
%
The terms are not unifiable due to the incompatibility at position 1.1
($\cst{c}$ vs.~$\cst{d}$).

@@

\noindent
where $\textsf{sk}$ is a fresh Skolem symbol,
$J$ is the largest subset of $\{1, \ldots, n$\} such that $x_j = y_j$ for every $j \in J$.
We denote the subsequence of $\tuple{x}{n}$ indexed by $J$ by $\tuple{x}{J}$.
Moreover, we require that
$x_i \neq y_i$, all variables in $\tuple{x}{K} \cdot \tuple{y}{K}$ are distinct,
where $K = \{1, \ldots, n\} \setminus J$,
and neither $\tuple{x}{K}$ nor $\tuple{y}{K}$ shares variables with $\tuple{x}{J}$.
For example, given $\cst{add} \; a \; b \noteq \cst{add} \; a \; b' \mathrel\lor a \eq b'$,
\infname{IR} can derive the existence of the inverse
$\cst{sk}_1$ characterized by $\cst{sk}_{1\!} \; (\cst{add} \; a \; b) \; a \eq b$.
%As a second example, from the clause
%$\cst{cons} \; x \; \mathit{xs} \noteq \cst{cons} \; y \; \mathit{ys}
%\mathrel\lor x \eq y$, we can derive
%$\cst{sk}_2 \; (\cst{cons} \; x \; \mathit{xs}) \eq x$.

@@

}
Configurations that are suitable for a wide range of problems have emerged over
time. One of them is the configuration that is most often chosen by E's \emph{auto}
mode. We call it \emph{boa} (``best of \emph{auto}'')\begin{conf}.\end{conf}\begin{rep}:
%
%%% TYPESETTING: TEMPORARILY DISABLED LEFTMARGIN (?)
\begin{description}[labelwidth=\widthof{Precedence generation:}]
  \item[\rm Term order:] KBO
  \item[\rm Weight generation:] \texttt{invfreqrank}
  \item[\rm Precedence generation:] \texttt{invfreq}
  \item[\rm Literal selection:] \texttt{SelectMLCAPP}
  \item[\rm Clause selection:] ~ \\
\verb|1.CRSWeight(SimulateSOS, 0.5, 100, 100, 100, 100, 1.5, 1.5, 1),| \\
\verb|4.CRSWeight(ConstPrio, 0.1, 100, 100, 100, 100, 1.5, 1.5, 1.5),| \\
\verb|1.FIFOWeight(PreferProcessed),| \\
\verb|1.CRSWeight(PreferNonGoals, 0.5, 100, 100, 100, 100, 1.5, 1.5, 1),| \\
\verb|4.Refinedweight(SimulateSOS, 3, 2, 2, 1.5, 2)|
\end{description}
%
The clause selection scheme consists of five queues, each of which is
specified by a weight function template. The prefixes $n$\texttt{.} next to
the template names indicate that the queue will be visited $n$ times in the round-robin scheme
before moving to the next one. The first argument to each
template is the clause priority scheme.

@@

There is another transformation performed by preprocessing that is
problematic, but for a different reason. \emph{Definition unfolding} is the
process of replacing equationally defined symbols with their definitions and
removing the defining equations. A \relax{definition} is a clause of the form
$\cst{f} \; \tuple{x}{m} \eq t$, where the variables $\tuple{x}{m}$ are distinct,
$\cst{f}$ does not occur in the right-hand side $t$, and $\Var(t) \subseteq
\{x_1,\dots,x_m\}$.
This transformation preserves unsatisfiability (provability) for first-~and
higher-order logic, but not for \lfhol{}, making Ehoh incomplete. The reason
is that by removing the definitional clause, we also remove a symbol $\cst{f}$
that otherwise could be used to instantiate a higher-order quantifier.
For example, the clause set
$\{\cst{f}\;x \eq x{,}\; \cst{f}\;(y\;\cst{a}) \noteq \cst{a}\}$ is unsatisfiable,
whereas $\{y\;\cst{a} \noteq \cst{a}\}$ is satisfiable in \lfhol{}. (In full
higher-order logic, the second clause set would be unsatisfiable thanks to
the $\{y \mapsto \lambda x.\; x\}$ instance and $\beta$-conversion.)
%
For the moment, we have simply disabled definition unfolding in Ehoh. We will
enable it again once we have added support for $\lambda$-terms.

@@

  \begin{sloppypar}
  Traditionally, most higher-order provers were designed from the ground up
  to target higher-order logic. Two
  exceptions are Otter-$\lambda$ by Beeson \cite{mb-04-lam-logic} and
  Zipperposition by Cruanes et al.\ %, Vukmirovi\'c, and colleagues
  \cite{sc-15-simon-phd,bbtvw-21-sup-lam}. \hbox{Otter-$\lambda$}
  adds $\lambda$-terms and second-order unification to
  the superposition prover Otter\begin{rep} \cite{mcc-03-otter}\end{rep}.
  % The approach is pragmatic. , with little emphasis on completeness.
  Zipperposition, also based on superposition, was extended
  to Bool\-e\-an-free higher-order logic by Bentkamp et al.\
  \cite{bbtvw-21-sup-lam}. Its performance is a far cry from E's, but
  it is easier to modify.
  Vukmirovi\'c et al.\ also used it to test and evaluate higher-order unification
  procedures \cite{vbn-21-unif} and Boolean reasoning
  \cite{vn-20-bools}.
  Zipperposition now includes Ehoh as a backend in a cooperative
  architecture.
  Finally, there is work by the developers of Vampire
  % br-20-full-sup-w-combs
  \cite{br-19-restricted-unif} and of the SMT
  (satisfiability modulo theories) solvers CVC4 and veriT
  \cite{brotb-19-ho-smt} to extend their provers to higher-order logic.
  \end{sloppypar}

@@

  An alternative to all of the above is to reduce higher-order logic to first-order
  logic via a translation. Robinson \cite{ar-70-hol}
  outlined this approach decades before tools such as
  \begin{conf}Sledgehammer \cite{pb-12-sh} and
  HOLyHammer \cite{ku-15-holyhammer}\end{conf}%
  \begin{rep}Miz$\mathbb{AR}$ \cite{urs-13-atp},
  Sledgehammer \cite{pb-12-sh},
  HOLyHammer \cite{ku-15-holyhammer}, and
  CoqHammer \cite{ck-18-coqhammer}\end{rep}
  popularized it in proof assistants.
  In addition to performing an applicative encoding, such translations must
  eliminate the \hbox{$\lambda$-expressions}\begin{rep}
  \cite{mp-08-trans,lc-16-lam-trans}\end{rep} and encode the type
  information\begin{rep} \cite{bkpu-16-hammering-types}\end{rep}.
  %
  \begin{rep}%
  In practice, on problems with a large first-order component, translations
  perform very well compared with the existing native provers
  \cite{ns-13-leo2sh}. Largely thanks to Sledgehammer, Isabelle often came
  in close second at CASC, even defeating Satallax in 2012
  \cite{gs-13-cascj6}.
  \end{rep}

>>>>

An easy but inefficient way to implement unification and matching for \lfhol{}
is to apply the applicative encoding (Sect.~\ref{sec:ehoh:logic}), perform
first-order unification or matching, and decode the resulting
substitution. To avoid overhead, we generalize the first-order unification
and matching procedures to operate directly on \lfhol{} terms.

@@

The key difference between $\Gfpf$ and $\GfpfRTL$ concerns applied variables.
Given the sample positions $(\varepsilon, 2, 1)$,
the fingerprint of $x$ is $(\AA, \BB, \BB)$ as before, whereas
the fingerprint of $x \; \cst{c}$ is $(\AA, \BB, \cst{c}).$
%
As another example,
let $(\varepsilon, 2,\allowbreak 1,\allowbreak 2.2,\allowbreak 2.1,\allowbreak 1.2, 1.1)$
be the sample positions,
and let $s = x \; (\cst{f} \; \cst{b} \; \cst{c})$ and $t = \cst{g}\; \cst{a}\; (y \; \cst{d}).$
Their fingerprints are
$\FpRTL(s) = (\AA, \BB, \cst{f}, \BB, \BB, \cst{b}, \cst{c})$ and
$\FpRTL(t) = (\cst{g}, \cst{a}, \AA, \NN, \NN, \BB, \cst{d})$.
%
The terms are not unifiable due to the incompatibility at position 1.1
($\cst{c}$ vs.~$\cst{d}$).

@@

\noindent
where $\textsf{sk}$ is a fresh Skolem symbol, and
$J$ is the largest subset of $\{1, \ldots, n$\} such that $x_j = y_j$ for every $j \in J$.
We denote the subsequence of $\tuple{x}{n}$ indexed by $J$ by $\tuple{x}{J}$.
Moreover, we require that
$x_i \neq y_i$, all variables in $\tuple{x}{K} \cdot \tuple{y}{K}$ are distinct,
where $K = \{1, \ldots, n\} \setminus J$,
and neither $\tuple{x}{K}$ nor $\tuple{y}{K}$ shares variables with $\tuple{x}{J}$.
For example, given $\cst{add} \; a \; b \noteq \cst{add} \; a \; b' \mathrel\lor a \eq b'$,
\infname{IR} can derive the existence of the inverse
$\cst{sk}_1$ characterized by $\cst{sk}_{1\!} \; (\cst{add} \; a \; b) \; a \eq b$.
%As a second example, from the clause
%$\cst{cons} \; x \; \mathit{xs} \noteq \cst{cons} \; y \; \mathit{ys}
%\mathrel\lor x \eq y$, we can derive
%$\cst{sk}_2 \; (\cst{cons} \; x \; \mathit{xs}) \eq x$.

@@

}
Configurations that are suitable for a wide range of problems have emerged over
time. One of them is the configuration that is most often chosen by E's \emph{auto}
mode. We call it \emph{boa} (``best of \emph{auto}'')\begin{conf}.\end{conf}\begin{rep}:
%
%%% TYPESETTING: TEMPORARILY DISABLED LEFTMARGIN (?)
\begin{description}[labelwidth=\widthof{Precedence generation:}]
  \item[\rm Term order:] KBO
  \item[\rm Weight generation:] \texttt{invfreqrank}
  \item[\rm Precedence generation:] \texttt{invfreq}
  \item[\rm Literal selection:] \texttt{SelectMLCAPP}
  \item[\rm Clause selection:] ~ \\
\verb|1.CRSWeight(SimulateSOS, 0.5, 100, 100, 100, 100, 1.5, 1.5, 1),| \\
\verb|4.CRSWeight(ConstPrio, 0.1, 100, 100, 100, 100, 1.5, 1.5, 1.5),| \\
\verb|1.FIFOWeight(PreferProcessed),| \\
\verb|1.CRSWeight(PreferNonGoals, 0.5, 100, 100, 100, 100, 1.5, 1.5, 1),| \\
\verb|4.Refinedweight(SimulateSOS, 3, 2, 2, 1.5, 2)|
\end{description}
%
The clause selection scheme consists of five queues, each of which is
specified by a weight function template. The prefixes $n$\texttt{.}\ next to
the template names indicate that the queue will be visited $n$ times in the round-robin scheme
before moving to the next one. The first argument to each
template is the clause priority scheme.

@@

There is another transformation performed by preprocessing that is
problematic, but for a different reason. \emph{Definition unfolding} is the
process of replacing equationally defined symbols with their definitions and
removing the defining equations. A \relax{definition} is a clause of the form
$\cst{f} \; \tuple{x}{m} \eq t$, where the variables $\tuple{x}{m}$ are distinct,
$\cst{f}$ does not occur in the right-hand side $t$, and $\Var(t) \subseteq
\{x_1,\dots,x_m\}$.
This transformation preserves unsatisfiability (provability) for first-~and
higher-order logic, but not for \lfhol{}, making Ehoh incomplete. The reason
is that by removing the definitional clause, we also remove a symbol $\cst{f}$
that otherwise could be used to instantiate a higher-order quantifier.
For example, the clause set
$\{\cst{f}\;x \eq x{,}\; \cst{f}\;(y\;\cst{a}) \noteq \cst{a}\}$ is unsatisfiable,
whereas $\{y\;\cst{a} \noteq \cst{a}\}$ is satisfiable in \lfhol{}. (In full
higher-order logic, the second clause set would be unsatisfiable thanks to
the $\{y \mapsto \lambda x.\; x\}$ instance and $\beta$-conversion.)
%
For the moment, we have simply disabled definition unfolding in Ehoh. We have
not measured the effect of this choice, but we conjecture it is not substantial.
Namely, appropriately chosen term order will result in a similar outcome without
removing the defined symbol. We will enable definition unfolding again once we
have added support for $\lambda$-terms.

@@

  \begin{sloppypar}
  Traditionally, most higher-order provers were designed from the ground up
  to target higher-order logic. Two
  exceptions are Otter-$\lambda$ by Beeson \cite{mb-04-lam-logic} and
  Zipperposition by Cruanes et al.\ %, Vukmirovi\'c, and colleagues
  \cite{sc-15-simon-phd,bbtvw-21-sup-lam}. \hbox{Otter-$\lambda$}
  adds $\lambda$-terms and second-order unification to
  the superposition prover Otter\begin{rep} \cite{mcc-03-otter}\end{rep}.
  % The approach is pragmatic. , with little emphasis on completeness.
  Zipperposition, also based on superposition, was extended
  to Bool\-e\-an-free higher-order logic by Bentkamp et al.\
  \cite{bbtvw-21-sup-lam}. Its performance is a far cry from E's, but
  it is easier to modify.
  Vukmirovi\'c et al.\ also used it to test and evaluate higher-order unification
  procedures \cite{vbn-21-unif} and Boolean reasoning
  \cite{vn-20-bools}.
  Zipperposition now includes Ehoh as a backend in a cooperative
  architecture.
  Finally, there is work by the developers of Vampire
  % br-20-full-sup-w-combs
  \cite{br-19-restricted-unif} and of the SMT solvers CVC4 and veriT
  \cite{brotb-19-ho-smt} to extend their provers to higher-order logic.
  \end{sloppypar}

@@

  An alternative to all of the above is to reduce higher-order logic to first-order
  logic via a translation. Robinson \cite{ar-70-hol}
  outlined this approach decades before tools such as
  \begin{conf}Sledgehammer \cite{pb-12-sh} and
  HOLyHammer \cite{ku-15-holyhammer}\end{conf}%
  \begin{rep}Miz$\mathbb{AR}$ \cite{urs-13-atp},
  Sledgehammer \cite{pb-12-sh},
  HOLyHammer \cite{ku-15-holyhammer}, and
  CoqHammer \cite{ck-18-coqhammer}\end{rep}
  popularized it in proof assistants.
  In addition to performing an applicative encoding, such translations must
  eliminate the \hbox{$\lambda$-expressions}\begin{rep}
  \cite{mp-08-trans,lc-16-lam-trans}\end{rep} and encode the type
  information\begin{rep} \cite{bkpu-16-hammering-types}\end{rep}.
  %
  \begin{rep}%
  In practice, on problems with a large first-order component, translations
  perform very well compared with the existing native provers
  \cite{ns-13-leo2sh}. Largely thanks to Sledgehammer, Isabelle often came
  in a close second at CASC, even defeating Satallax in 2012
  \cite{gs-13-cascj6}.
  \end{rep}

### 2022-03-08 introduction/introduction.tex
more

Various researchers also experimented with the idea of extending a first-order
prover to higher-order logic. Beeson \cite{mb-04-lam-logic} implemented second-order
unification and added $\lambda$-terms to one of the best superposition
provers at the time, Otter~\cite{mcc-03-otter}. Cruanes extended Zipperposition
\cite{sc-15-simon-phd} with arithmetic, induction, and rudimentary support for
higher-order logic. Bhayat and Reger implemented a complete superposition
calculus  for higher-order logic, based on combinators  in Vampire
\cite{br-20-full-sup-w-combs}. Authors of SMT solvers followed a similar path:
they extended cvc4~\cite{cbetal-11-cvc4} and veriT~\cite{bodf-09-veriT} to
support higher-order logic \cite{brotb-19-ho-smt}.

>>>>

Various researchers also experimented with the idea of extending a first-order
prover to higher-order logic. Beeson \cite{mb-04-lam-logic} implemented second-order
unification and added $\lambda$-terms to one of the best superposition
provers at the time, Otter~\cite{mcc-03-otter}. Cruanes extended Zipperposition
\cite{sc-15-simon-phd} with arithmetic, induction, and rudimentary support for
higher-order logic. Bhayat and Reger implemented a complete superposition
calculus  for higher-order logic, based on combinators  in Vampire
\cite{br-20-full-sup-w-combs}. Authors of SMT (satisfiability modulo theories) solvers followed a similar path:
they extended cvc4~\cite{cbetal-11-cvc4} and veriT~\cite{bodf-09-veriT} to
support higher-order logic \cite{brotb-19-ho-smt}.

### 2022-03-08 pre/pre.tex
more

Before spelling out the inference rules of standard, first-order superposition
let us introduce some helpful notation, following Schulz \cite{ss-02-brainiac}.
With $t|_p$ we denote subterm of $t$ at position $p$ and with $t[p \leftarrow
t']$ replacement of this subterm with $t'$. Let \selfun{} be a selection function,
$\succ$ a reduction ordering, $C = l \llor C'$ a clause, and $\sigma$ a
substitution. We say that \sigmacl{l} is eligible for resolution if: (1) nothing
is selected by \selfun{} and \sigmacl{l} is $\succ$-maximal within the literals in
\sigmacl{C} or (2) \selfun{} selected some literals and \sigmacl{l} is the maximal
within either positive or negative selected literals. We say \sigmacl{l} is
eligible for paramodulation if it is positive, nothing is selected and
\sigmacl{l} is the maximal literal within the literals of \sigmacl{C}. With
$\mgu(s,t)$ we denote the most general unifier of $s$ and $t$. Using a horizontal
line to separate premises and the conclusion, four inference rules of
superposition are as follows:

@@

\begin{tabular}{m{\dimexpr 0.36\linewidth-2\tabcolsep}m{\dimexpr 0.57\linewidth-2\tabcolsep}}
    \multicolumn{2}{l}{{\bf Superposition into positive literals} (SP)} \\[\jot]
    $\namedinference{SP}{s \eq t \llor C \qquad u \eq v \llor D}{\sigmacl{u[p \leftarrow t] \eq v \llor C \llor D}}$ 
        & where $\sigma = \mgu(u|_p, s), \sigma(t) \not\succ \sigma(s), \sigma(v) \not\succ \sigma(u)$,
        $\sigmacl{s \eq t}$ is eligible for paramodulation, just like $\sigmacl{u \eq v}$ and $u|_p$ is not a variable
\end{tabular}

@@

\medskip
Let us give some intuition behind how these rules are applied in practice. The
\infname{ER} rule is used to replace equality with equality within a clause.
Consider the formula $\iforall x.\,  x \eq \cst{2} \iimplies \cst{even}(x)$. In
clausal form it is $X \noteq \cst{2} \llor \cst{even}(X)$. Applying the
\infname{ER} rule effectively applies the precondition $X \eq \cst{2}$ to the
conclusion $\cst{even}(X)$ to obtain $\cst{even}(\cst{2})$. The \infname{EF}
rule can be used to find instances of the clause that can be further simplified.
Consider the clause $\cst{f}(X) \eq \cst{a} \llor \cst{f}(Y) \eq \cst{a} \llor
\cst{p}(X, Y)$. Applying \infname{EF} to its first two literals yields $\cst{a}
\noteq \cst{a} \llor \cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$. This clause
then further simplifies to $\cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ using \infname{ER}.

Lastly, superposition rules are used to simulate what mathematicians do when
they ``replace equals by equals''. Left premise intuitively states that an
equation $s \eq t$ holds under condition $C$ (or more precisely, its negation).
Similarly, the right premise asserts that a (dis)equation holds under a
condition. Superposition rule simply concatenates the conditions for both
premises and replaces equals by equals in two main inference terms. For example
given clauses $\neg \cst{even}(X) \llor \cst{f}(X) \eq \cst{a}$ and $\neg
\cst{even}(\cst{b}) \llor \cst{g}(\cst{f}(Y)) \eq \cst{b}$, superposition between their
last literals results in $\neg
\cst{even}(X) \llor \neg \cst{even}(\cst{b}) \llor \cst{g}(\cst{a}) \eq
\cst{b}$.

>>>>

Before spelling out the inference rules of standard, first-order superposition
let us introduce some helpful notation, following Schulz \cite{ss-02-brainiac}.
With $t|_p$ we denote the subterm of $t$ at position $p$ and with $t[p \leftarrow
t']$ replacement of this subterm with $t'$. Let \selfun{} be a selection function,
$\succ$ a reduction ordering, $C = l \llor C'$ a clause, and $\sigma$ a
substitution. We say that \sigmacl{l} is eligible for resolution if: (1) nothing
is selected by \selfun{} and \sigmacl{l} is $\succ$-maximal within the literals in
\sigmacl{C} or (2) \selfun{} selected some literals and \sigmacl{l} is the maximal
within either positive or negative selected literals. We say \sigmacl{l} is
eligible for paramodulation if it is positive, nothing is selected and
\sigmacl{l} is the maximal literal within the literals of \sigmacl{C}. With
$\mgu(s,t)$ we denote the most general unifier of $s$ and $t$. Using a horizontal
line to separate premises and the conclusion, four inference rules of
superposition are as follows:

@@

\begin{tabular}{m{\dimexpr 0.36\linewidth-2\tabcolsep}m{\dimexpr 0.57\linewidth-2\tabcolsep}}
    \multicolumn{2}{l}{{\bf Superposition into positive literals} (SP)} \\[\jot]
    $\namedinference{SP}{s \eq t \llor C \qquad u \eq v \llor D}{\sigmacl{u[p \leftarrow t] \eq v \llor C \llor D}}$ 
        & where $\sigma = \mgu(u|_p, s), \sigma(t) \not\succ \sigma(s), \sigma(v) \not\succ \sigma(u)$,
        $\sigmacl{s \eq t}$ and $\sigmacl{u \eq v}$ are eligible for paramodulation, and $u|_p$ is not a variable
\end{tabular}

@@

\medskip
Let us give some intuition behind how these rules are applied in practice. The
\infname{ER} rule is used to replace equality with equality within a clause.
Consider the formula $\iforall x.\,  x \eq \cst{2} \iimplies \cst{even}(x)$. In
clausal form it is $X \noteq \cst{2} \llor \cst{even}(X)$. Applying the
\infname{ER} rule effectively applies the precondition $X \eq \cst{2}$ to the
conclusion $\cst{even}(X)$ to obtain $\cst{even}(\cst{2})$. The \infname{EF}
rule can be used to find instances of the clause that can be further simplified.
Consider the clause $\cst{f}(X) \eq \cst{a} \llor \cst{f}(Y) \eq \cst{a} \llor
\cst{p}(X, Y)$. Applying \infname{EF} to its first two literals yields $\cst{a}
\noteq \cst{a} \llor \cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ (with $\sigma = \{ Y \mapsto X \}$). This clause
then further simplifies to $\cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ using \infname{ER}.

Lastly, the rules \infname{SP} and \infname{SN} are used to simulate what mathematicians do when
they ``replace equals by equals''. The left premise intuitively states that an
equation $s \eq t$ holds under condition $C$ (or more precisely, its negation).
Similarly, the right premise asserts that a (dis)equation holds under a
condition. The superposition rule simply concatenates the conditions for both
premises and replaces equals by equals in two main inference terms. For example
given clauses $\neg \cst{even}(X) \llor \cst{f}(X) \eq \cst{a}$ and $\neg
\cst{even}(\cst{b}) \llor \cst{g}(\cst{f}(Y)) \eq \cst{b}$, superposition between their
last literals results in $\neg
\cst{even}(X) \llor \neg \cst{even}(\cst{b}) \llor \cst{g}(\cst{a}) \eq
\cst{b}$.
